### 2020-06-26 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet fix

Finding the whole set of efficient decisions may require a lot of computations and is often deemed redundant for the decision-maker. As a result, approaches that allow obtaining only particular limited sets of effective decisions are used in practice more often.

@@

A note worthy approach to solving MCO problems is lexicographic optimization methods. Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}. When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value. This sequence is simple and clear and allows using many effective scalar optimization algorithms. Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}. At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker. The $\varepsilon$-constraint scalarization approach can be used to expand the set of effective decisions that can be considered by the decision-maker. In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}. Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of restrictions is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem. Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties. A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems). The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for effective decisions, when the order of criteria importance can also change dynamically. It is necessary to note that the presence of variability in problem statement for the search of effective decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations. Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}. In \cite{c1} a combined approach is proposed to the application of scalarization criteria and additional constraints of the $\varepsilon$-constraint method.

@@

Such unimprovable decisions are called Pareto efficient or optimal.

@@

\begin{equation}\label{eq:9}
\begin{matrix}
f^*_1 =\min_{y \in D}{f_1(y)},  \\
f^*_i =\min_{y \in D}{f_i(y)},  f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s, \\
P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s.
\end{matrix}
\end{equation}

@@

\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function in the point of the global minimum should be compared with the values of the function in all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points,
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on eof the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function in these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

@@

\begin{itemize}
	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
	\item Intensive search information obtained in the process of computations for solving multiple global optimization problems.
\end{itemize}

@@

The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function in these points), that is
\begin{equation}\label{eq:11}
y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
\end{equation}
where $k$ is the number of the global search iterations performed, $G_k$ is the decisive rule of the global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ iteration of the global search. The computational complexity of the decisive rule can be significantly simplified by reducing the optimization problems by the Peano space-filling curves or evolvents $y(x)$ that uniquely and continuously map the interval [0,1] to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}). As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:7}) is reduced to a one-dimensional problem:
\begin{equation}\label{eq:12}
f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots, f_s(y(x))) \to min_{lex},  x \in [0,1].
\end{equation}

@@

\begin{equation}\label{eq:24}
h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
\end{equation}

If $h_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken. For the initial values of these parameters, $M^0 = H_l^0 = 1$, $z_l^0 = 0$, $1 \leq i \leq m+1$, can be taken.

@@

\begin{theorem}
If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the values $h_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
\begin{equation} \label{eq:29}
r_\nu h_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
\end{equation}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
\end{theorem}

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of computational experiments.

@@

With a complete scheme for calculating criteria values, i.e. when at the iteration points $x_i$, $1 \leq i \leq k$ the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant. In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$ and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations. The results of computational experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.

@@

\begin{theorem}
Let the next solved optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previously solved problem $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1]
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l h_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, L_i \text{from} (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:37}
\Psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k, 
\end{equation}
(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $h_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).
\end{theorem}

@@

\begin{theorem}
Let the optimization problem $\phi(x)$ of the computational scheme (\ref{eq:9}) be solved before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials for the interval $[a,b] \subset [0,1]$ is reached
\begin{equation}\label{eq:39}
p_{ab}=\frac{n_{ab}}{b-a}
\end{equation}
where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$. Then, when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
\begin{equation}\label{eq:40}
(3m/2\Delta)(b-a), m \text{from} (\ref{eq:35}),
\end{equation}
will be limited by the value
\begin{equation}\label{eq:41}
n_{ab}' \leq (3m/2\Delta - p_{ab})(b-a)
\end{equation}
if the condition is met
\begin{equation}\label{eq:42}
\psi(x) \geq \psi(x^*)+\Delta, x \in [a,b].
\end{equation}
\end{theorem}

@@

\section{Results of Computational Experiments}
\label{sec:5}

The computational experiments were carried out on the Lobachevsky supercomputer at the Lobachevsky State University of Nizhny Novgorod (the CentOS 6.4 operating system and the SLURM workload manager). A supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. Each CPU is 8-core, which means 16 CPU cores are available on the node. The Intel C ++ 14.0.2 compiler was used to obtain an executable program code. The numerical experiments were performed using the Globalizer system \cite{c32}.

@@

In this article we assessed the efficiency of the MCO methods using two widely used indicators \cite{c35,c36}:
\begin{itemize}
	\item The hypervolume index (HV). This indicator evaluates the approximation of the Pareto set in terms of completeness; a higher value corresponds to a more complete coverage of the Pareto set.
	\item The distribution uniformity index (DU). This indicator evaluates the uniformity of Pareto set coverage; a lower value corresponds to a more uniform coverage of the Pareto set.
\end{itemize}

@@

Another important issue of the efficiency evaluation consists in the choice of benchmarks, which will be used to measure the efficiency of the methods applied. Unfortunately, there is no unified approach to this issue either. In many studies, the efficiency of the developed methods is demonstrated by the example of a limited set of test optimization problems. Researchers repeatedly attempt to resolve this issue by forming quite a wide set of test optimization problems, which can be used to evaluate the efficiency of the MCO methods. For instance, in \cite{c53} authors propose a flexible toolkit for constructing well-designed MCO problems. In \cite{c54} a set of 24 test optimization functions is given, which are widely used for evaluating the MCO methods. In \cite{c55} a set of 100 test problems is considered. In \cite{c56} an approach is proposed, in which a set of efficiency parameters of the MCO problems is formed from the well-known scalar optimization functions. This allows forming an almost endless set of MCO problems. At the same time, the set of the MCO test problems, in which the efficiency criteria would be multiextremal functions, is rather limited. Also, the proposed sets usually consist of single examples of the MCO problems, which makes it difficult to obtain any reliable estimates of the method efficiency.

@@

In the first series of computational experiments, the MAGCS algorithm was compared with several multicriteria optimization algorithms for solving the two-criteria problem proposed in \cite{c35}:
\begin{equation}\label{eq:43}
f_1 (y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0\leq y_1,y_2 \leq 1.
\end{equation}

@@

The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto domain did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).

In the second series of experiments, the two-criteria two-dimensional MCO problems were solved. The solution was carried out with the method accuracy $\varepsilon = 0.025$ and the reliability parameters $r_1 = r_2 = 5.6$. The results of computational experiments are presented in Table \ref{tab:2}.

@@

The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto domain after a substantially smaller total number of iterations of the global search (the PI columns). It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto region with better HV and DU indices.

For a better visual representation, Figure \ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information. The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.

@@

It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subtasks of the same MCO task.

@@

The third set of computational experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations. The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig. \ref{fig:4}--\ref{fig:6}. These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCO subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).

@@

In conclusion, it can be noted that the approach described here is promising, and requires further study. It is necessary to continue numerical experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems. It is also necessary to evaluate the possibility of parallel computing using high-performance supercomputer systems.

@@

\bibitem{c44} Charles Audet, Gilles Savard, and Walid Zghal: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)

@@

\bibitem{c56} Dimo Brockhoff, Tea Tusar, Anne Auger, Nikolaus Hansen: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites. (2019) https://arxiv.org/abs/1604.00359 

>>>>

Finding the whole set of efficient decisions may require a lot of computations and is often deemed redundant for the decision-maker. As a result, approaches that allow obtaining only particular limited sets of efficient decisions are used in practice more often.

@@

A note worthy approach to solving MCO problems is lexicographic optimization methods. Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}. When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value. This sequence is simple and clear and allows using many efficient scalar optimization algorithms. Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}. At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker. The $\varepsilon$-constraint scalarization approach can be used to expand the set of efficient decisions that can be considered by the decision-maker. In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}. Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of constraints is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem. Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties. A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems). The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for efficient decisions, when the order of criteria importance can also change dynamically. It is necessary to note that the presence of variability in problem statement for the search of efficient decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations. Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}. In \cite{c1} a combined approach is proposed to the application of scalarization criteria and additional constraints of the $\varepsilon$-constraint method.

@@

Such unimprovable decisions are called Pareto efficient or Pareto optimal.

@@

\begin{equation}\label{eq:9}
\begin{matrix}
f^*_1 =\min_{y \in D}{f_1(y)},  \\
f^*_i =\min_{y \in D}{f_i(y)},  f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s, \\
P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < s.
\end{matrix}
\end{equation}

@@

\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points,
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

@@

\begin{itemize}
	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
	\item Intensive use of the search information obtained in the process of computations for solving multiple global optimization problems.
\end{itemize}

@@

The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function at these points), that is
\begin{equation}\label{eq:11}
y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
\end{equation}
where $k$ is the number of the global search iterations performed, $G_k$ is the decisive rule of the global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ iteration of the global search. The computational complexity of the decisive rule can be significantly simplified by reducing the optimization problems by the Peano space-filling curves or evolvents $y(x)$ that uniquely and continuously map the interval [0,1] to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}). As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:7}) is reduced to a one-dimensional problem:
\begin{equation}\label{eq:12}
f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots, f_s(y(x))) \to min_{lex},  x \in [0,1].
\end{equation}

@@

\begin{equation}\label{eq:24}
H^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
\end{equation}

If $H_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $H_l^k=1$ is taken. For the initial values of these parameters, $M^0 = H_l^0 = 1$, $z_l^0 = 0$, $1 \leq i \leq m+1$, can be taken.

@@

\begin{theorem}
If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
\begin{equation} \label{eq:29}
r_\nu H^k_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
\end{equation}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
\end{theorem}

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

@@

With a complete scheme for calculating criteria values, i.e. when at the iteration points $x_i$, $1 \leq i \leq k$ the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant. In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$ and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations. The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.

@@

\begin{theorem}
Let the next solved optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previously solved problem $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1]
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k, 
\end{equation}
\end{theorem}(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).

@@

\begin{theorem}
Let the optimization problem $\phi(x)$ of the computational scheme (\ref{eq:9}) be solved before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials for the interval $[a,b] \subset [0,1]$ is reached
\begin{equation}\label{eq:39}
p_{ab}=\frac{n_{ab}}{b-a}
\end{equation}
where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$. Then, when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
\begin{equation}\label{eq:40}
(3m/2\Delta)(b-a), m \; \text{from} \;  (\ref{eq:35}),
\end{equation}
will be limited by the value
\begin{equation}\label{eq:41}
n_{ab}' \leq (3m/2\Delta - p_{ab})(b-a)
\end{equation}
if the condition is met
\begin{equation}\label{eq:42}
\psi(x) \geq \psi(x^*)+\Delta, x \in [a,b].
\end{equation}
\end{theorem}

@@

\section{Results of Numerical Experiments}
\label{sec:5}

The numerical experiments were carried out on the Lobachevsky supercomputer at the Lobachevsky State University of Nizhny Novgorod (the CentOS 6.4 operating system and the SLURM workload manager). A supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. Each CPU is 8-core, which means 16 CPU cores are available on the node. The Intel C ++ 14.0.2 compiler was used to obtain an executable program code. The numerical experiments were performed using the Globalizer system \cite{c32}.

@@

In this paper we assessed the efficiency of the MCO methods using two widely used indicators \cite{c35,c36}:
\begin{itemize}
	\item The hypervolume index (HV). This indicator evaluates the approximation of the Pareto set in terms of completeness; a higher value corresponds to a more complete coverage of the Pareto set.
	\item The distribution uniformity index (DU). This indicator evaluates the uniformity of Pareto set coverage; a lower value corresponds to a more uniform coverage of the Pareto set.
\end{itemize}

@@

Another important issue of the efficiency evaluation consists in the choice of benchmarks, which will be used to measure the efficiency of the methods applied. Unfortunately, there is no unified approach to this issue either. In many studies, the efficiency of the developed methods is demonstrated by the example of a limited set of test optimization problems. Researchers repeatedly attempt to resolve this issue by forming quite a wide set of test optimization problems, which can be used to evaluate the efficiency of the MCO methods. For instance, in \cite{c53} authors propose a flexible toolkit for constructing well-designed MCO problems. In \cite{c54} a set of 24 test optimization functions is given, which are widely used for evaluating the MCO methods. In \cite{c55} a set of 100 test problems is considered. In \cite{c56} an approach is proposed, in which a set of efficiency criteria of the MCO problems is formed from the well-known scalar optimization functions. This allows forming an almost endless set of MCO problems. At the same time, the set of the MCO test problems, in which the efficiency criteria would be multiextremal functions, is rather limited. Also, the proposed sets usually consist of single examples of the MCO problems, which makes it difficult to obtain any reliable estimates of the method efficiency.

@@

In the first series of numerical experiments, the MAGCS algorithm was compared with several multicriteria optimization algorithms for solving the two-criteria problem proposed in \cite{c35}:
\begin{equation}\label{eq:43}
f_1 (y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0\leq y_1,y_2 \leq 1.
\end{equation}

@@

The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).

In the second series of experiments, the two-criteria two-dimensional MCO problems were solved. The solution was carried out with the method accuracy $\varepsilon = 0.025$ and the reliability parameters $r_1 = r_2 = 5.6$. The results of numerical experiments are presented in Table \ref{tab:2}.

@@

The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto set after a substantially smaller total number of iterations of the global search (the PI columns). It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto set with better HV and DU indices.

For a better visual representation, Fig. \ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information. The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.

@@

It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problems.

@@

The third set of numerical experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations. The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig. \ref{fig:4}--\ref{fig:6}. These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCO subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).

@@

In conclusion, it can be noted that the approach described here is promising, and requires further studies. It is necessary to continue numerical experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems. It is also necessary to evaluate the possibility of parallel computing using high-performance supercomputer systems.

@@

\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)

@@

\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites. (2019) https://arxiv.org/abs/1604.00359 
